---
title: "Ataxia RNAseq - Leafcutter junction category distributions"
author: 
- name: "Guillermo Rocamora Pérez"
  affiliation: UCL - Version 1.3
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  bookdown::html_document2:
    figure_caption: yes
    code_folding: show
    theme: paper
    highlight: haddock
    df_print: paged
    toc: true
    toc depth: 3
    toc_float: true
    number_sections: true
  md_document:
    variant: markdown_github
    toc: true
    number_sections: true
always_allow_html: true
---

```{r setup, include = F}
# Initial setup ----
## Required libraries
shhh <- suppressPackageStartupMessages
shhh(library(here))
shhh(library(tidyverse))
shhh(library(foreach))
shhh(library(patchwork))
shhh(library(GenomeInfoDb))
shhh(library(GenomicRanges))

## Define main path
main_path <- "/home/grocamora/Core_Projects/Ataxia_Splicing_Analysis/"

## Load helper functions
source(file.path(main_path, "R/Leafcutter_introns.R"))
source(file.path(main_path, "R/hf_graph_and_themes.R"))

## Set options
options(dplyr.summarise.inform = FALSE)
options(lifecycle_verbosity = "warning")
knitr::opts_chunk$set(echo = F, warning = F, message = F, out.width="85%", fig.align = "center", dpi = 300)
```

```{r load-theme, echo = F, results = "asis"}
## Custom styles for the output html
cat('
<style type="text/css">
.dataTables_scrollHeadInner{
  width:100% !important;
}
.dataTables_scrollHeadInner table{
  width:100% !important;
}
/*
.code-folding-btn {
  display: none;
}
*/
h3, .h3 {
  font-size: 22px!important;
}
h4, .h4 {
  font-size: 18px!important;
}
h5, .h5 {
  font-size: 16px!important;
}
body{
  font-size: 13px;
}
.tocify-subheader {
    text-indent: 15px;
    display: none;
    font-size: 12px;
}
</style>')

# Dictionary to rename the comparisons in the plot
intron_type_labels <- c("annotated" = "Annotated", "novel_donor" = "Novel Donor", 
                        "novel_acceptor" = "Novel Acceptor", 
                        "novel_combo" = "Novel Combo", "novel_exon_skip" = "Novel Exon Skip", 
                        "unannotated" = "Unannotated")

intron_type_levels <- c("annotated", "novel_acceptor", "novel_donor", 
                        "novel_exon_skip", "novel_combo", "unannotated")
```

# Background
> Aim: update the studies carried out by Regina H. Reynolds about introns' annotation distributions. Original source material: [Leafcutter: intron annotation and visualisation](https://rhreynolds.github.io/LBD-seq-bulk-analyses/quantification_splicing/Leafcutter_intron_annotation.html)

This report is meant to be used as a template. As such, the different sections will include instructions on how to modify the graphic outputs and not descriptions about the results themselves. To run this script, it is necessary to source the file at `/home/grocamora/Core_Projects/Ataxia_Splicing_Analysis/R/Leafcutter_introns.R` with needed helper functions.

## Changelog {.tabset -}

### v1.3 {-}

* Changed the reference transcriptome to Gencode v38.

### v1.2 {-}

* Changed the reference transcriptome to Gencode v39.

# Annotate introns

To annotate the junctions, we need to set the following variables:

* `overwrite_results`: whether to overwrite previously annotated results. If set to `FALSE`, it will look for the file in disk (given a path) and load the results. If set to `TRUE`, it will always annotate the introns.

* `use_strand`: whether to use the strand information for the Leafcutter clusters' ID. 

* `gtf_path`: path to the reference transcriptome to annotate the junctions. In this report, GENCODE v41 is employed.

* `results_path`: path to here to store the results. Also needed to specify the path to figures and where to store the annotated introns.

```{r, echo = T}
## Script Parameters
overwrite_results <- F
use_strand <- F
gtf_path <- "/home/grocamora/RytenLab-Research/Additional_files/GENCODE/gencode.v38.annotation.gtf"

results_path <- file.path(main_path, "results/leafcutter_annotated/")
figure_path <- file.path(results_path, "figures/")
annotated_path <- file.path(results_path, "annotated_gencode38.rds")

dir.create(results_path, showWarnings = F, recursive = T)
dir.create(figure_path, showWarnings = F, recursive = T)
```

## Loading leafcutter results

Next, the leafcutter results are loaded from disk. Three different studies will be presented in this report: overall (Level 1), subtypes (Level 2) and diagnoses (Level 3). Types 2 and 3 will be further split by tissues.

```{r, echo = T}
path_to_results <- file.path(main_path, "data/leafcutter_results/")
leafcutter_list <-  setNames(list(
  readRDS(file.path(path_to_results, "case_v_control_leafcutter_ds.Rds")),
  readRDS(file.path(path_to_results, "subtypes_v_control_leafcutter_ds.Rds")),
  readRDS(file.path(path_to_results, "diagnoses_v_control_leafcutter_ds.Rds"))),
  c("overall", "subtypes", "diagnoses"))
```

## Annotation process

The junctions are annotated using the function `junction_annot` from David Zhang's package [`dasper`](http://www.bioconductor.org/packages/release/bioc/html/dasper.html) version 1.7.0. 

```{r, echo = T}
# Path to where the annotated intron list will be stored in disk.
annotated_path <- file.path(results_path, "annotated_gencode38.rds")

# If file already exists, do not execute the annotation process.
if(!file.exists(annotated_path)){
  # Load the reference transcriptome
  TxDb_ref <- dasper:::ref_load(gtf_path)
  
  # Some reference transcriptomes (i.e. Gencode) use the characters "chr" before
  # the seqnames. To bring consistency, the "seqlevelsStyle" is set to "NCBI".
  #GenomeInfoDb::seqlevelsStyle(TxDb_ref) <- c("NCBI")
  
  annotated <- leafcutter_list %>% purrr::map(function(x){
    # Extract the successful clusters
    success_cluster <- x$cluster_significance %>%
      dplyr::filter(status == "Success") %>% 
      dplyr::pull(cluster) %>% 
      unique()
    
    # Transform the data from leafcutter to a valid GRanges object to run the annotation
    annotated_introns <- x$intron_usage %>%
      tidyr::separate(col = "intron", into = c("chr", "intron_start", "intron_end", "cluster_id"), sep = ":", remove = F) %>%
      dplyr::mutate(cluster_id = str_c(chr, ":", cluster_id)) %>%
      dplyr::filter(cluster_id %in% success_cluster) %>%
      dplyr::select(-chr, -intron_start, -intron_end, cluster_id) %>%
      convert_leafcutter(use_strand = use_strand, seqlevelsStyle = GenomeInfoDb::seqlevelsStyle(TxDb_ref)) %>%
      dasper::junction_annot(ref = TxDb_ref)
    
    # Once the introns are annotated, we remove all the annotated introns that
    # are associated to more than one gene.
    annotated_introns <- removeAmbiguousGenesLeafcutter(annotated_introns)
    
    return(annotated_introns)
  })
  
  # Save the annotated list in disk.
  annotated %>% saveRDS(annotated_path)
}else{
  annotated <- readRDS(annotated_path)
}
```

It is important to use a consistent transcriptome `seqlevelsStyle` in order to properly match the extracted junctions. By default, `NCBI` style is recommended (no `chr` before the `seqnames`).

## Proportion of annotated introns {.tabset}

We use the function `printTypeTable` to print a table with the proportion of significant junctions by category. Results by tissue are shown when expanded below the main table. The significant column denotes the number of introns by annotation type that were found significantly differentially spliced. [Original Source](https://rhreynolds.github.io/LBD-seq-bulk-analyses/quantification_splicing/Leafcutter_intron_annotation.html#annotating-introns).

```{r, echo = T, eval = T}
#' Prints the proportion of significant junctions by type
#'
#' Given the list of annotated junctions and the list of leafcutter results, this
#' function extract a given index of the lists (corresponding to different
#' studies) and prints a table with the proportion of significantly
#' differentially spliced junction by type.
#'
#' @param annotated list of annotated junctions. Each element corresponds to a specific study.
#' @param leafcutter_list list of leafcutter results. Each element corresponds to a specific study.
#' @param level index or name of the specific study in both the "annotated" and "leafcutter_list" arguments.
#' @param use_strand boolean to specify whether to use the strand in the clusters' ID.
#' @param deltapsi_filter boolean to specify whether to use a deltaPSI filter of |dPSI| >= 0.1
#' @param tissue character vector, if provided, the comparisons employed must contain the keyword.
#'
#' @export
printTypeTable <- function(annotated, leafcutter_list, level, use_strand = F, deltapsi_filter = F, tissue = NULL){
  # Extract the significant comparisons and clusters
  significant_clusters <- leafcutter_list[[level]]$significant_clusters_0.05_filter %>%
    dplyr::mutate(cluster_id = str_c(chr, ":", cluster)) %>% 
    removeClusterStrand(use_strand, input_name = "cluster_id") %>%
    {if(deltapsi_filter) dplyr::filter(., abs(deltapsi) >= 0.1) else .} %>%
    dplyr::distinct(comparison, cluster_id)
  
  # Conver the annotated GRanges object to a tibble.
  annotated_df <- annotated[[level]] %>%
    tibble::as_tibble()
  
  # If a tissue is provided, filter the comparisons to contain the input tissue.
  if(!is.null(tissue)){
    significant_clusters <- significant_clusters %>%
      dplyr::filter(grepl(tissue, comparison))
    annotated_df <- annotated_df %>%
      dplyr::filter(grepl(tissue, comparison))
  }
  
  # Extract the number of junctions by type that are significant.
  significant_counts <- annotated_df %>%
    dplyr::inner_join(significant_clusters, by = c("comparison", "cluster_id")) %>%
    dplyr::group_by(type) %>% # (Optional) Add comparison/tissue here.
    dplyr::summarise(significant = n())
  
  # Prints the proportion of significant junctions by type.
  output_table <- annotated_df %>%
    dplyr::group_by(type) %>% # (Optional) Add comparison/tissue here.
    dplyr::summarise(all = n()) %>% 
    dplyr::inner_join(significant_counts, by = "type") %>%
    dplyr::mutate(proportion = (significant/all) %>% signif(2)) %>%
    dplyr::arrange(-all) %>% 
    `colnames<-`(c("Junction Type", "Count Junctions", "Significant Junctions", "Proportion"))
  
  output_table %>% 
    kableExtra::kbl(booktabs = T, linesep = "") %>%
    kableExtra::kable_classic(full_width = F, "hover", "striped", html_font = "Cambria", font_size = 14) %>%
    kableExtra::row_spec(0, bold = T, font_size = 16)
}
```

### Level 1 (Type) {.tabset -}

#### Lenient {-}

```{r}
printTypeTable(annotated, leafcutter_list, level = "overall", use_strand = use_strand)
```

<details><summary>Information by tissue</summary>

Cerebellum:

```{r}
printTypeTable(annotated, leafcutter_list, level = "overall", use_strand = use_strand, tissue = "Cerebellum")
```

Frontal Cortex:

```{r}
printTypeTable(annotated, leafcutter_list, level = "overall", use_strand = use_strand, tissue = "Frontal")
```

</details>

#### Stringent {-}

```{r}
printTypeTable(annotated, leafcutter_list, level = "overall", use_strand = use_strand, deltapsi_filter = T)
```

<details><summary>Information by tissue</summary>

Cerebellum:

```{r}
printTypeTable(annotated, leafcutter_list, level = "overall", use_strand = use_strand, deltapsi_filter = T, tissue = "Cerebellum")
```

Frontal Cortex:

```{r}
printTypeTable(annotated, leafcutter_list, level = "overall", use_strand = use_strand, deltapsi_filter = T, tissue = "Frontal")
```

</details>

### Level 2 (AtaxiaSubtype) {.tabset -}

#### Lenient {-}

```{r}
printTypeTable(annotated, leafcutter_list, level = "subtypes", use_strand = use_strand)
```

<details><summary>Information by tissue</summary>

Cerebellum:

```{r}
printTypeTable(annotated, leafcutter_list, level = "subtypes", use_strand = use_strand, tissue = "Cerebellum")
```

Frontal Cortex:

```{r}
printTypeTable(annotated, leafcutter_list, level = "subtypes", use_strand = use_strand, tissue = "Frontal")
```

</details>

#### Stringent {-}

```{r}
printTypeTable(annotated, leafcutter_list, level = "subtypes", use_strand = use_strand, deltapsi_filter = T)
```

<details><summary>Information by tissue</summary>

Cerebellum:

```{r}
printTypeTable(annotated, leafcutter_list, level = "subtypes", use_strand = use_strand, deltapsi_filter = T, tissue = "Cerebellum")
```

Frontal Cortex:

```{r}
printTypeTable(annotated, leafcutter_list, level = "subtypes", use_strand = use_strand, deltapsi_filter = T, tissue = "Frontal")
```

</details>

### Level 3 (Diagnosis) {.tabset -}

#### Lenient {-}

```{r}
printTypeTable(annotated, leafcutter_list, level = "diagnoses", use_strand = use_strand)
```

<details><summary>Information by tissue</summary>

Cerebellum:

```{r}
printTypeTable(annotated, leafcutter_list, level = "diagnoses", use_strand = use_strand, tissue = "Cerebellum")
```

Frontal Cortex:

```{r}
printTypeTable(annotated, leafcutter_list, level = "diagnoses", use_strand = use_strand, tissue = "Frontal")
```

</details>

#### Stringent {-}

```{r}
printTypeTable(annotated, leafcutter_list, level = "diagnoses", use_strand = use_strand, deltapsi_filter = T)
```

<details><summary>Information by tissue</summary>

Cerebellum:

```{r}
printTypeTable(annotated, leafcutter_list, level = "diagnoses", use_strand = use_strand, deltapsi_filter = T, tissue = "Cerebellum")
```

Frontal Cortex:

```{r}
printTypeTable(annotated, leafcutter_list, level = "diagnoses", use_strand = use_strand, deltapsi_filter = T, tissue = "Frontal")
```

</details>

# Visualization of overlapping events

Different visualizations will be presented replicating Regina's results, but adapting the code to remove depecrated functions and adding the option to use non-stranded clusters.

## Proportion of annotation types within each comparison {.tabset}

Visualization of the proportion of junction categories by comparison and tissue. We use the function `plotProportionOfAnnotation` with the parameters explained next. Relevant to notice the addition of the `tissue` parameter that can split the comparisons by keywords in their name (e.g. if we specify `Cerebellum` in the tissue parameter, only comparison that include the keyword `Cerebellum` will be considered). [Original Source](https://rhreynolds.github.io/LBD-seq-bulk-analyses/quantification_splicing/Leafcutter_intron_annotation.html#number-of-overlapping-introns)

```{r, echo = T, eval = T}
#' Plots the proportion of junction categories
#'
#' Given the list of annotated junctions and the list of leafcutter results,
#' this function extract a given index of the lists (corresponding to different
#' studies) and plots the proportion of each category.
#'
#' Additional parameters include the option to use the cluster strand or to
#' divide by tissue.
#'
#' @param annotated list of annotated junctions. Each element corresponds to a
#'   specific study.
#' @param leafcutter_list list of leafcutter results. Each element corresponds
#'   to a specific study.
#' @param level index or name of the specific study in both the "annotated" and "leafcutter_list" arguments.
#' @param use_strand boolean to specify whether to use the strand in the
#'   clusters' ID.
#' @param tissue character vector with the specific tissue to study. The tissue
#'   name must be contained in the comparison name.
#'
#' @return
plotProportionOfAnnotation <- function(annotated, leafcutter_list, level, use_strand = F, tissue = NULL){
  # Get the successful comparisons and clusters
  success_cluster <- leafcutter_list[[level]]$cluster_significance %>%
    # If tissue is provided, it filter the comparisons by tissue
    {if(!is.null(tissue)) dplyr::filter(., grepl(tissue, comparison)) else .} %>%
    dplyr::filter(status == "Success") %>%
    dplyr::distinct(comparison, cluster) %>%
    removeClusterStrand(use_strand)
  
  # Extract the introns with successful comparison and cluster
  success <- annotated[[level]] %>%
    tibble::as_tibble() %>%
    dplyr::inner_join(success_cluster, by = c("comparison", "cluster_id"))
  
  # Get introns with significant comparison and cluster
  significant <- extractSignificantIntrons(annotated[[level]], leafcutter_list[[level]], use_strand, tissue)
  
  # Plots will be stored in a list 
  intron_list <- list(success, significant)
  plot_list <- vector(mode = "list", 2)
  
  for(j in 1:length(intron_list)){
    # Calculate the proportion of each category by comparison
    count <- intron_list[[j]] %>% 
      dplyr::filter(!type == "ambig_gene") %>% 
      dplyr::group_by(comparison, type) %>% 
      dplyr::summarise(n = n()) %>% 
      dplyr::mutate(prop = n/sum(n)) %>% 
      dplyr::ungroup()
    
    plot_list[[j]] <- count %>% 
      dplyr::mutate(type = type %>% factor(levels = intron_type_levels %>% rev()),
                    #comparison = comparison_labels[comparison],
                    comparison = as_factor(comparison)) %>% 
      ggplot(aes(x = comparison, y = prop, fill = type), colour = "black") +
      geom_col(color = "black", width= 0.6) +
      # ggrepel::geom_text_repel(aes(label = paste0(round(prop, 2)*100, "%"), x = comparison, group = type), 
      #                          position = ggpp::position_stacknudge(x = 0.4, direction = "split"),
      #                          size = 3.5) +
      # ggrepel::geom_text_repel(aes(label = round(prop, 2), x = comparison, group = type), direction = "x", force= 10) +
      labs(x = "", y = "Proportion") +
      scale_y_continuous(expand = expansion(mult = c(0, 0.05))) + 
      scale_x_discrete(expand = expansion(add = 0.5)) + 
      scale_fill_manual(name = "Acceptor/donor annotation", 
                        values = rev(c("#3C5488", "#E64B35", "#00A087", "#4DBBD5", "#7E6148", "grey"))) +
      custom_gg_theme + 
      theme(axis.text.x = ggplot2::element_text(color = "black", size = 8, angle = 90, hjust = 0.5)) +
      guides(fill = guide_legend(reverse = T))
  }
  
  # Arrange both plots in two columns
  ggpubr::ggarrange(plotlist = plot_list,
                    labels = c("a", "b"),
                    common.legend = TRUE, legend = "top") %>%
    ## If tissue is provided, add a title (remove next line to ignore title)
    {if(!is.null(tissue)) ggpubr::annotate_figure(., top = ggpubr::text_grob(paste0("Tissue: ", tissue), size = 16)) else .}
}
```

### Level 1 (Type) {-}

```{r, fig.height=5, fig.width=8, fig.cap="Proportion of (a) all successfully tested introns and (b) the subset of differentially spliced introns coloured by annotation type."}
plotProportionOfAnnotation(annotated, leafcutter_list, level = "overall", use_strand = use_strand)
```

### Level 2 (AtaxiaSubtype) {-}

#### Cerebellum {-}

```{r, fig.height=5, fig.width=8, fig.cap="Proportion of (a) all successfully tested introns and (b) the subset of differentially spliced introns coloured by annotation type. Results only for Cerebellum tissue."}
plotProportionOfAnnotation(annotated, leafcutter_list, level = "subtypes", use_strand = use_strand, tissue = "Cerebellum")
```

#### Frontal Cortex {-}

```{r, fig.height=5, fig.width=8, fig.cap="Proportion of (a) all successfully tested introns and (b) the subset of differentially spliced introns coloured by annotation type. Results only for Frontal Cortex tissue."}
plotProportionOfAnnotation(annotated, leafcutter_list, level = "subtypes", use_strand = use_strand, tissue = "Frontal")
```

### Level 3 (Diagnosis) {-}

#### Cerebellum {-}

```{r, fig.height=5, fig.width=8, fig.cap="Proportion of (a) all successfully tested introns and (b) the subset of differentially spliced introns coloured by annotation type. Results only for Cerebellum tissue."}
plotProportionOfAnnotation(annotated, leafcutter_list, level = "diagnoses", use_strand = use_strand, tissue = "Cerebellum")
```

#### Frontal Cortex {-}

```{r, fig.height=5, fig.width=8, fig.cap="Proportion of (a) all successfully tested introns and (b) the subset of differentially spliced introns coloured by annotation type. Results only for Frontal Cortex tissue."}
plotProportionOfAnnotation(annotated, leafcutter_list, level = "diagnoses", use_strand = use_strand, tissue = "Frontal")
```

## Number of overlapping introns {.tabset}

In the following plots, we measure the number of differentially spliced introns across overlaps. We use the function `plotNumberOverlappingIntrons`, with the same arguments as before. [Original Source](https://rhreynolds.github.io/LBD-seq-bulk-analyses/quantification_splicing/Leafcutter_intron_annotation.html#number-of-overlapping-introns)

```{r, echo = T, eval = T}
#' Plot the number of overlapping significant junctions
#'
#' First, we obtain the annotated introns from significant clusters and
#' comparisons. Then, we remove the junction category "ambig_gene" and count the
#' number of entries by intron ID and category.
#'
#' @param annotated list of annotated junctions. Each element corresponds to a
#'   specific study.
#' @param leafcutter_list list of leafcutter results. Each element corresponds
#'   to a specific study.
#' @param level index or name of the specific study in both the "annotated" and "leafcutter_list" arguments.
#' @param use_strand boolean to specify whether to use the strand in the
#'   clusters' ID.
#' @param tissue character vector with the specific tissue to study. The tissue
#'   name must be contained in the comparison name.
#'
#' @return
#' @export
plotNumberOverlappingIntrons <- function(annotated, leafcutter_list, level, use_strand = T, tissue = NULL){
  # Get introns with significant comparison and cluster
  significant <- extractSignificantIntrons(annotated[[level]], leafcutter_list[[level]], use_strand, tissue)
  
  # Find number of overlapped introns between the comparisons
  number_ovelaps <- significant %>% 
    dplyr::filter(!type == "ambig_gene") %>% 
    dplyr::mutate(unique_id = str_c(cluster_id, ":", start, ":", end),
                  type = type %>% factor(levels = intron_type_levels)) %>% 
    dplyr::group_by(unique_id, type) %>% 
    dplyr::mutate(n_overlaps = n()) %>%
    dplyr::ungroup() %>%
    dplyr::mutate(n_overlaps = factor(n_overlaps, levels = 1:length(unique(comparison)))) %>%
    dplyr::distinct(unique_id, type, n_overlaps)
  
  # Plot the results
  ggplot(number_ovelaps, aes(x = n_overlaps, fill = type)) +
    geom_bar(position = position_dodge2(preserve = "single", padding = 0), color = "black") +
    facet_wrap(~ type, labeller = labeller(type = intron_type_labels)) +
    labs(x = "Number of overlaps") +
    scale_y_continuous(expand = expansion(mult = c(0, 0.05))) + 
    scale_fill_manual(name = "Acceptor/donor annotation", 
                      values = c("#3C5488", "#E64B35", "#00A087", "#4DBBD5", "#7E6148", "grey"),
                      labels = intron_type_labels) +
    custom_gg_theme +
    theme(axis.text.x = ggplot2::element_text(color = "black", size = 8, angle = 0, hjust = 0.5)) +
    ## If tissue is provided, add a title (remove next line to ignore title)
    {if(!is.null(tissue)) ggtitle(paste0("Tissue: ", tissue))}
}
```

### Level 1 (Type) {-}

```{r, fig.height=4.5, fig.width=7.2, fig.cap = "Number of differentially spliced introns across overlaps."}
plotNumberOverlappingIntrons(annotated, leafcutter_list, level = "overall", use_strand = use_strand)
```

### Level 2 (AtaxiaSubtype) {-}

#### Cerebellum {-}

```{r, fig.height=5, fig.width=7.2, fig.cap = "Number of differentially spliced introns across overlaps."}
plotNumberOverlappingIntrons(annotated, leafcutter_list, level = "subtypes", use_strand = use_strand, tissue = "Cerebellum")
```

#### Frontal Cortex {-}

```{r, fig.height=5, fig.width=7.2, fig.cap = "Number of differentially spliced introns across overlaps."}
plotNumberOverlappingIntrons(annotated, leafcutter_list, level = "subtypes", use_strand = use_strand, tissue = "Frontal")
```

### Level 3 (Diagnosis) {-}

#### Cerebellum {-}

```{r, fig.height=5, fig.width=7.2, fig.cap = "Number of differentially spliced introns across overlaps."}
plotNumberOverlappingIntrons(annotated, leafcutter_list, level = "diagnoses", use_strand = use_strand, tissue = "Cerebellum")
```

#### Frontal Cortex {-}

```{r, fig.height=5, fig.width=7.2, fig.cap = "Number of differentially spliced introns across overlaps."}
plotNumberOverlappingIntrons(annotated, leafcutter_list, level = "diagnoses", use_strand = use_strand, tissue = "Frontal")
```

## Distribution of annotation types across comparisons {.tabset}

We can also plot by comparisons the proportion of the number of overlapping introns. [Original Source](https://rhreynolds.github.io/LBD-seq-bulk-analyses/quantification_splicing/Leafcutter_intron_annotation.html#distribution-of-annotation-types-across-comparisons).

```{r, echo = T, eval = T}
#' Plot the proportion of overlapping significant junctions by comparison
#'
#' First, we obtain the annotated introns from significant clusters and
#' comparisons. Then, we remove the junction category "ambig_gene" and count the
#' number of entries by intron ID and category. The proportion of each category
#' is calculated for each number of overlaps.
#'
#' @param annotated list of annotated junctions. Each element corresponds to a
#'   specific study.
#' @param leafcutter_list list of leafcutter results. Each element corresponds
#'   to a specific study.
#' @param level index or name of the specific study in both the "annotated" and "leafcutter_list" arguments.
#' @param use_strand boolean to specify whether to use the strand in the
#'   clusters' ID.
#' @param tissue character vector with the specific tissue to study. The tissue
#'   name must be contained in the comparison name.
#'
#' @return
#' @export
plotDistributionAnnotatedTypes <- function(annotated, leafcutter_list, level, use_strand = F, tissue = NULL){
  # Get introns with significant comparison and cluster
  significant <- extractSignificantIntrons(annotated[[level]], leafcutter_list[[level]], use_strand, tissue)
  
  # Find number of overlapped introns between the comparisons
  number_ovelaps <- significant %>% 
    dplyr::filter(!type == "ambig_gene") %>% 
    dplyr::mutate(unique_id = str_c(cluster_id, ":", start, ":", end),
                  type = type %>% factor(levels = intron_type_levels)) %>% 
    dplyr::group_by(unique_id, type) %>% 
    dplyr::mutate(n_overlaps = n() %>% as_factor()) %>%
    dplyr::ungroup() %>%
    dplyr::mutate(n_overlaps = factor(n_overlaps, levels = 1:length(unique(comparison))))
  
  # Find the proportion of each category for a given number of overlaps
  number_overlaps_proportion <- number_ovelaps %>% 
    dplyr::distinct(comparison, unique_id, n_overlaps, type) %>% 
    dplyr::group_by(comparison, n_overlaps, type) %>% 
    dplyr::summarise(n = n()) %>% 
    dplyr::group_by(comparison) %>% 
    dplyr::mutate(prop = n/sum(n))
  
  # Plot the results
  ggplot(number_overlaps_proportion,
         aes(x = n_overlaps, y = prop, fill = type)) +
    geom_col(position = position_dodge2(preserve = "single", padding = 0), color = "black") +
    facet_grid(type ~ comparison, scales = "free_y", labeller = labeller(type = intron_type_labels)) +
    labs(y = "Proportion of differentially spliced introns") +
    scale_y_continuous(expand = expansion(mult = c(0, 0.05))) + 
    scale_fill_manual(name = "Acceptor/donor annotation", 
                      values = c("#3C5488", "#E64B35", "#00A087", "#4DBBD5", "#7E6148", "grey"),
                      labels = intron_type_labels) +
    custom_gg_theme +
    theme(axis.text.x = ggplot2::element_text(color = "black", size = 8, angle = 0, hjust = 0.5)) +
    {if(!is.null(tissue)) ggtitle(paste0("Tissue: ", tissue))}
}
```

### Level 1 (Type) {-}

```{r, fig.height=8, fig.width=7.2, fig.cap = "Proportion of differentially spliced introns per overlap across annotation types and comparisons. Proportions were calculated per comparison, by dividing the number of distinct events in each annotation type by the total number across a comparison."}
plotDistributionAnnotatedTypes(annotated, leafcutter_list, level = "overall", use_strand = use_strand)
```

### Level 2 (AtaxiaSubtype) {-}

#### Cerebellum {-}

```{r, fig.height=8, fig.width=9, fig.cap = "Proportion of differentially spliced introns per overlap across annotation types and comparisons. Proportions were calculated per comparison, by dividing the number of distinct events in each annotation type by the total number across a comparison."}
plotDistributionAnnotatedTypes(annotated, leafcutter_list, level = "subtypes", use_strand = use_strand, tissue = "Cerebellum")
```

#### Frontal Cortex {-}

```{r, fig.height=8, fig.width=9, fig.cap = "Proportion of differentially spliced introns per overlap across annotation types and comparisons. Proportions were calculated per comparison, by dividing the number of distinct events in each annotation type by the total number across a comparison."}
plotDistributionAnnotatedTypes(annotated, leafcutter_list, level = "subtypes", use_strand = use_strand, tissue = "Frontal")
```

### Level 3 (Diagnosis) {-}

#### Cerebellum {-}

```{r, fig.height=8, fig.width=9, fig.cap = "Proportion of differentially spliced introns per overlap across annotation types and comparisons. Proportions were calculated per comparison, by dividing the number of distinct events in each annotation type by the total number across a comparison."}
plotDistributionAnnotatedTypes(annotated, leafcutter_list, level = "diagnoses", use_strand = use_strand, tissue = "Cerebellum")
```

#### Frontal Cortex {-}

```{r, fig.height=8, fig.width=9, fig.cap = "Proportion of differentially spliced introns per overlap across annotation types and comparisons. Proportions were calculated per comparison, by dividing the number of distinct events in each annotation type by the total number across a comparison."}
plotDistributionAnnotatedTypes(annotated, leafcutter_list, level = "diagnoses", use_strand = use_strand, tissue = "Frontal")
```

## Proportion of clusters containing only annotated introns {.tabset}

Lastly, we also represent the proportion of significant cluster in which all the introns are annotated introns. [Original Source](https://rhreynolds.github.io/LBD-seq-bulk-analyses/quantification_splicing/Leafcutter_intron_annotation.html#proportion-of-clusters-containing-only-annotated-introns)

```{r, echo = T, eval = T}
#' Plot the proportion significant clusters containing only annotated introns
#'
#' Obtain the introns from significant clusters and comparisons and removes the
#' junction category "ambig_gene". The proportion of annotated introns in each
#' cluster and comparison is calculated and represented.
#'
#' @param annotated list of annotated junctions. Each element corresponds to a
#'   specific study.
#' @param leafcutter_list list of leafcutter results. Each element corresponds
#'   to a specific study.
#' @param level index or name of the specific study in both the "annotated" and "leafcutter_list" arguments.
#' @param use_strand boolean to specify whether to use the strand in the
#'   clusters' ID.
#' @param tissue character vector with the specific tissue to study. The tissue
#'   name must be contained in the comparison name.
#'
#' @return
#' @export
plotDistributionOnlyAnnotatedTypes <- function(annotated, leafcutter_list, level, use_strand = F, tissue = NULL){
  # Get introns with significant comparison and cluster
  significant <- extractSignificantIntrons(annotated[[level]], leafcutter_list[[level]], use_strand, tissue)
  
  # Find number of overlapped introns between the comparisons
  number_ovelaps <- significant %>% 
    dplyr::filter(!type == "ambig_gene") %>% 
    dplyr::mutate(unique_id = str_c(cluster_id, ":", start, ":", end),
                  type = type %>% factor(levels = intron_type_levels)) %>% 
    dplyr::group_by(unique_id, type) %>% 
    dplyr::mutate(n_overlaps = n() %>% as_factor()) %>%
    dplyr::ungroup()
  
  # Count the proportion of clusters with all their introns categorized as
  # annotated
  cluster_annot_count <- number_ovelaps %>% 
    dplyr::group_by(comparison, cluster_id) %>% 
    dplyr::summarise(introns_all_annotated = all(in_ref)) %>% 
    dplyr::group_by(comparison, introns_all_annotated) %>% 
    dplyr::summarise(n = n()) %>% 
    dplyr::mutate(prop = n/sum(n)) %>% 
    dplyr::ungroup()
  
  # Plot the results
  cluster_annot_count %>% 
    dplyr::mutate(comparison = as_factor(comparison)) %>% 
    ggplot(aes(x = comparison, y = prop, fill = introns_all_annotated)) +
    geom_col(color = "black") +
    geom_text(aes(label = paste0(round(prop, 2)*100, "%")), position = position_stack(vjust = 0.5), 
              size = 4, color = "black") +
    scale_y_continuous(expand = expansion(mult = c(0, 0.05))) + 
    scale_fill_manual(name = "All introns annotated", 
                      values = c("#E64B35", "#3C5488")) +
    labs(y = "Proportion of clusters") +
    custom_gg_theme +
    {if(!is.null(tissue)) ggtitle(paste0("Tissue: ", tissue))}
}
```

### Level 1 (Type) {-}

```{r, fig.height=5, fig.width=7.2, fig.cap="Proportion of differentially spliced clusters containing introns that are all fully annotated."}
plotDistributionOnlyAnnotatedTypes(annotated, leafcutter_list, level = "overall", use_strand = use_strand)
```

### Level 2 (AtaxiaSubtype) {-}

```{r, fig.height=6.5, fig.width=7.2, fig.cap="Proportion of differentially spliced clusters containing introns that are all fully annotated."}
plotDistributionOnlyAnnotatedTypes(annotated, leafcutter_list, level = "subtypes", use_strand = use_strand) +
  theme(axis.text.x = ggplot2::element_text(color = "black", size = 8, angle = 90, hjust = 0.5))
```

### Level 3 (Diagnosis) {-}

```{r, fig.height=6.5, fig.width=7.2, fig.cap="Proportion of differentially spliced clusters containing introns that are all fully annotated."}
plotDistributionOnlyAnnotatedTypes(annotated, leafcutter_list, level = "diagnoses", use_strand = use_strand) +
  theme(axis.text.x = ggplot2::element_text(color = "black", size = 8, angle = 90, hjust = 0.5))
```