---
title: "RBP Result"
output: 
  html_document:
    number_sections: true
---

```{r eval = F, include = F}
knit_output <- "html"
base_level <- 1
experiment_path <- here::here("projects/Frontal-Case_vs_Control-RIN4/")
tissue <- "Frontal (RIN > 4)"

shhh <- suppressPackageStartupMessages
shhh(library(here))
shhh(library(doFuture))
shhh(library(foreach))
shhh(library(tidyverse))
shhh(library(ggpubr))
options(dplyr.summarise.inform = FALSE)
options(lifecycle_verbosity = "warning")
options(scipen=999)

source(here::here("Helper_Functions/hf_Subsampling.R"))
source(here::here("Helper_Functions/hf_Analysis.R"))

knitr::opts_chunk$set(echo = F, warning = F, message = F, out.width="85%", fig.align = "center", dpi = 300)

## Load metadata
metadata_path <- here::here("metadata/metadata.csv")

metadata <- readr::read_delim(metadata_path, show_col_types = FALSE) %>% 
  dplyr::rowwise() %>%
  dplyr::mutate(Individual_ID = stringr::str_split(ID_anon, "_", simplify = T)[1]) %>%
  dplyr::filter(!(Diagnosis %in% c("CANVAS", "AIFM1"))) %>%
  extractReadDepthMultiQC()


## Load common novel and common intron
common_introns <- readRDS(paste0(experiment_path, "common_introns.rds"))
common_novel <- readRDS(paste0(experiment_path, "common_novel.rds"))

## Theme
custom_gg_theme <- theme(plot.title = element_text(size = 11, face = "bold"),
                        panel.border = ggplot2::element_rect(colour = "black", fill = NA, linewidth = 1),
                        axis.text.x = ggplot2::element_text(color = "black", size = 8, angle = 0, hjust = 0.5),
                        axis.text.y = ggplot2::element_text(color = "black", size = 8),
                        axis.title.x = ggplot2::element_text(face = "bold", size = 11, margin=margin(5,0,0,0)),
                        axis.title.y = ggplot2::element_text(face = "bold", size = 11, margin=margin(0,10,0,0)),
                        panel.grid.minor = element_line(color = "#444444", linewidth  = 0.05, linetype = 2),
                        panel.grid.major.y = element_line(color = "#444444", linewidth  = 0.05, linetype = 2),
                        panel.grid.major.x = element_line(color = "#444444", linewidth  = 0.1),
                        panel.background = element_rect(fill = "#FBFBFB"),
                        legend.title = element_text(size=12),
                        legend.text = element_text(size=10),
                        legend.position = "top",
                        legend.key = element_rect(color="black"),
                        legend.key.size = unit(1, 'lines'),
                        strip.text.x = element_text(color = "black", face = "bold", size = 9),
                        strip.background = element_rect(color="black", linewidth=1, linetype="solid"),
                        strip.text.y = element_text(color = "black", face = "bold", size = 9),
                        plot.margin = margin(0.5, 0.5, 0.2, 0.5, "cm"))

custom_gg_theme_subtitle <- custom_gg_theme + 
  theme(plot.subtitle = element_text(size=12, vjust = 1, color="black", margin = margin(b = -20, t = -5)),
        plot.title = element_text(size = 15, face = "bold"),
        legend.margin=margin(b = -5),
        legend.justification = "right",
        legend.title = element_text(size=11),
        legend.text = element_text(size=10))

novel_labels <- c("novel_donor" = "Novel Donor", "novel_acceptor" = "Novel Acceptor")
novel_mini_labels <- c("donor" = "Novel Donor", "acceptor" = "Novel Acceptor")
sequence_labels <- c("acceptor exon" = "Acceptor exon",
                     "acceptor intron" = "Acceptor intron",
                     "donor exon" = "Donor exon",
                     "donor intron" = "Donor intron")
delta_MES_labels <- c("delta MES 5'ss" = "Delta MES 5'ss", "delta MES 3'ss" = "Delta MES 3'ss")
```

```{r include = F}
## Functions to run Rmd template properly
library(glue)
sprintf_transformer <- function(text, envir) {
  m <- regexpr(":.+$", text)
  if (m != -1) {
    format <- substring(regmatches(text, m), 2)
    regmatches(text, m) <- ""
    res <- eval(parse(text = text, keep.source = FALSE), envir)
    do.call(sprintf, list(glue("%{format}"), res))
  } else {
    eval(parse(text = text, keep.source = FALSE), envir)
  }
}
glue_fmt <- function(..., .envir = parent.frame()) {
  glue(..., .transformer = sprintf_transformer, .envir = .envir)
}

addTitles <- function(base_level, element_level){
  return(strrep("#", base_level + element_level))
}

print_df <- function(x,
                     style = "html",
                     limit = NULL, random = F, seed = NULL,
                     order = list(), autoWidth = FALSE, pageLength = 10, rownames = FALSE, compact = T,
                     rowGroup = NULL, hideRowGroup = T,
                     full_width = F){
  if(style == "html"){
    dt_class = "display cell-border nowrap"
    if(compact) dt_class = paste0(dt_class, " compact")

    if(is.null(rowGroup)){
      DT::datatable(x,
                    options = list(scrollX = TRUE,
                                   autoWidth = autoWidth,
                                   bLengthChange = F,
                                   pageLength = pageLength,
                                   order = order,
                                   dom = 't'),
                    class = dt_class,
                    rownames = rownames)
    }else{
      if(hideRowGroup){
        hideRow <- rowGroup
      }else{
        hideRow <- -1
      }
      
      DT::datatable(x, extensions = 'RowGroup',
                    options = list(rowGroup = list(dataSrc=c(rowGroup)),
                                   bLengthChange = F,
                                   columnDefs = list(list(visible=FALSE, targets=c(hideRow))),
                                   scrollX = TRUE,
                                   autoWidth = autoWidth,
                                   pageLength = pageLength,
                                   order = order,
                                   dom = 't'),
                    class = dt_class,
                    rownames = rownames)

    }

  }else{
    if(length(order) != 0){
      if(order[[2]] == "desc"){
        x <- x %>% dplyr::arrange(desc(.[[order[[1]]+1]]))
      }else{
        x <- x %>% dplyr::arrange(.[[order[[1]]+1]])
      }
    }

    if(!is.null(limit)){
      if(random){
        if(!is.null(seed)) set.seed(seed)
        x <- x %>% .[sample(1:nrow(.), limit), ]
      }else{
        x <- x %>% .[1:limit, ]
      }
    }

    x %>%
      kableExtra::kbl(booktabs = T, linesep = "") %>%
      kableExtra::kable_classic(full_width = full_width, "hover", "striped", html_font = "Cambria", font_size = 14) %>%
      kableExtra::row_spec(0, bold = T, font_size = 16)
  }
}
```

```{r results = "asis", echo = FALSE}
cat(addTitles(base_level, 1), " Distances (N) {-}\n", sep = "")
```

```{r echo=FALSE, message=F, warning = F, fig.width=7.2, fig.height = 5.5}
limit_bp = 30
ggplot(common_novel %>% 
         dplyr::filter(abs(distance) <= limit_bp) %>%
         dplyr::group_by(distance, cluster, novel_type) %>% dplyr::count()) + 
  geom_bar(aes(x = distance, y = n, fill = cluster), stat="identity", 
           position = "stack", color = "black", linewidth = 0.1, width = 1) +
  scale_x_continuous(expand = expansion(mult = c(0, 0)),
                     breaks = seq(-limit_bp, limit_bp, length.out = 7)) + 
  scale_y_continuous(expand = expansion(mult = c(0, 0.05))) + 
  viridis::scale_fill_viridis(discrete = T, begin = 0.15, end = 0.75, name = "") + 
  labs(x = "Distance", y = "Number of unique novel junctions") + 
  ggtitle(paste0("Distance to the annotated intron"),
          subtitle = paste0("Tissue: ", tissue)) + 
  ggforce::facet_col(vars(novel_type), labeller = labeller(novel_type = novel_labels)) + 
  custom_gg_theme_subtitle + theme(panel.spacing.x = unit(1, "lines"))
```

<details>
<summary>Expand details</summary>

* X-Axis: Distance between the novel splice site and the reference splice site.
* Y-Axis: Number of unique novel junctions. Bars are stacked.
* Colour: Cluster type as case vs. control.
* Facets: Novel junction category (Novel donor vs. Novel acceptor).
</details>

---

```{r echo=FALSE, message=F, warning = F, fig.width=7.2, fig.height=4}
limit_bp = 30
ggplot(common_novel %>%
         dplyr::filter(abs(distance) <= limit_bp) %>%
         dplyr::group_by(distance, cluster, novel_type) %>% dplyr::count()) + 
  geom_bar(aes(x = distance, y = n, fill = cluster), stat="identity", 
           position = "stack", color = "black", linewidth = 0.1, width = 1) +
  scale_x_continuous(expand = expansion(mult = c(0, 0)),
                     breaks = seq(-limit_bp, limit_bp, length.out = 7)) + 
  scale_y_continuous(expand = expansion(mult = c(0, 0.05))) + 
  viridis::scale_fill_viridis(discrete = T, begin = 0.15, end = 0.75, name = "") + 
  labs(x = "Distance", y = "Number of unique novel junctions") + 
  ggtitle(paste0("Distance to the annotated intron"),
          subtitle = paste0("Tissue: ", tissue)) + 
  facet_grid(novel_type ~ cluster, scales = "free_y",
             labeller = labeller(novel_type = novel_labels)) + 
  custom_gg_theme_subtitle + theme(panel.spacing.x = unit(1, "lines"))
```

<details>
<summary>Expand details</summary>

* X-Axis: Distance between the novel splice site and the reference splice site.
* Y-Axis: Number of unique novel junctions.
* Colour: Cluster type as case vs. control.
* Facets: Novel junction category (Novel donor vs. Novel acceptor) and cluster (same as colour).
</details>

---

```{r echo = FALSE, warning=F, fig.width=7.2, fig.height=5.5}
limit_bp = 30
common_novel %>%
  dplyr::filter(abs(distance) <= limit_bp) %>%
  dplyr::group_by(distance, novel_type, cluster) %>%
  dplyr::count() %>%
  tidyr::pivot_wider(id_cols = c("novel_type", "distance"), names_from = "cluster", values_from = "n") %>%
  dplyr::mutate(Case = replace_na(Case, 0),
                Control = replace_na(Control, 0)) %>%
  dplyr::mutate(diff = Case - Control) %>%
  ggplot() +
  geom_bar(aes(x = distance, y = diff, fill = "fill_color"), stat="identity", 
           position = "stack", color = "black", linewidth = 0.2, width = 1) +
  scale_x_continuous(expand = expansion(mult = c(0, 0)), 
                     breaks = seq(-limit_bp, limit_bp, length.out = 7)) + 
  scale_y_continuous(expand = expansion(mult = c(0.05, 0.05))) +
  scale_fill_manual(name = "", values = c("fill_color" = "#4394fd"),
                    labels = c("Case - Control"),
                    guide = guide_legend(override.aes = list(linewidth = .25))) +
  labs(x = "Distance", y = "Number of unique novel junctions") + 
  ggtitle(paste0("Distance difference to the annotated intron"),
          subtitle = paste0("Tissue: ", tissue)) + 
  ggforce::facet_col(vars(novel_type), labeller = labeller(novel_type = novel_labels)) +
  custom_gg_theme_subtitle
```

<details>
<summary>Expand details</summary>

* X-Axis: Distance between the novel splice site and the reference splice site.
* Y-Axis: Difference between the number of unique novel junctions from case samples and from control samples.
* Facets: Novel junction category (Novel donor vs. Novel acceptor).
</details>

---

```{r results = "asis", echo = FALSE}
cat(addTitles(base_level, 1), " Modulo 3 {.tabset -}\n", sep = "")
```

```{r echo = FALSE, fig.width=7.2, fig.height=5.5}
limit_bp = 30

modulo_df <- common_novel %>%
  dplyr::filter(abs(distance) <= 100) %>%
  tidyr::separate(novel_type, c("novel", "novel_type")) %>%
  dplyr::mutate(modulo = abs(distance) %% 3,
                sequence = ifelse(distance < 0, "intron", "exon"),
                novel_type = factor(novel_type, levels = c("acceptor", "donor"))) %>%
  dplyr::select(cluster, novel_type, sequence, distance, modulo)

ggplot(modulo_df) +
  geom_bar(aes(x = modulo, fill = cluster), stat = "count", position = "dodge", color = "black", linewidth = 0.5) +
  ggrepel::geom_text_repel(aes(x = modulo, label = after_stat(count), group = cluster), stat = "count",
                           size = 4.2, color = "white", bg.color = "black", bg.r = .1, fontface = "bold",
                           vjust = 1.5, position = position_dodge(width = 0.9), force = 5) +
  scale_x_continuous(limits = c(-0.5, 2.5), breaks = c(0, 1, 2), expand = c(0, 0)) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.05))) + 
  viridis::scale_fill_viridis(discrete = T, begin = 0.15, end = 0.75, name = "") +
  labs(x = "Modulo 3 of the distance", y = "Number of unique novel junctions") + 
  ggtitle(paste0("Mod.3 distance to the annotated intron"), subtitle = paste0("Tissue: ", tissue)) + 
  ggforce::facet_col(vars(novel_type), labeller = labeller(novel_type = novel_mini_labels)) +
  custom_gg_theme_subtitle + theme(legend.key = element_rect(color = "black", linewidth = 0.10))
```

<details>
<summary>Expand details</summary>

* X-Axis: Distance in modulo 3 between the novel splice site and the reference splice site. Filtered by $distance\leq 100\;bp$.
* Y-Axis: Number of unique novel junctions. Also represented as the text inside the bars.
* Colour: Cluster type as case vs. control.
* Facets: Novel junction category (Novel donor vs. Novel acceptor).
</details>

---

```{r echo = FALSE, fig.width=7.2, fig.height=5.5}
limit_bp = 30

modulo_group_df <- modulo_df %>%
  dplyr::group_by(cluster, novel_type, modulo) %>% 
  dplyr::summarise(n = n()) %>%
  dplyr::mutate(freq = (n / sum(n) * 100))

ggplot(modulo_group_df) +
  geom_bar(aes(x = modulo, y = freq, fill = cluster), stat = "identity", position = "dodge", color = "black", linewidth = 0.5) +
  ggrepel::geom_text_repel(aes(x = modulo, y = freq, label = paste0(round(freq, 1), " %"), group = cluster),
                           size = 4.2, color = "white", bg.color = "black", bg.r = .1, fontface = "bold",
                           vjust = 1.5, position = position_dodge(width = 0.9), force = 5) +
  scale_x_continuous(limits = c(-0.5, 2.5), breaks = c(0, 1, 2), expand = c(0, 0)) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.05))) + 
  viridis::scale_fill_viridis(discrete = T, begin = 0.15, end = 0.75, name = "") +
  labs(x = "Modulo 3 of the distance", y = "Percentage of novel junctions") + 
  ggtitle(paste0("Mod.3 Percentages: Cluster and Splice site"), subtitle = paste0("Tissue: ", tissue)) + 
  ggforce::facet_col(vars(novel_type), labeller = labeller(novel_type = novel_mini_labels)) +
  custom_gg_theme_subtitle + theme(legend.key = element_rect(color = "black", linewidth = 0.10))
```

<details>
<summary>Expand details</summary>

* X-Axis: Distance in modulo 3 between the novel splice site and the reference splice site. Filtered by $distance\leq 100\;bp$.
* Y-Axis: Percentage of each modulo 3 value with respect to cluster (case and control) and novel type (donor and acceptor).
* Colour: Cluster type as case vs. control.
* Facets: Novel junction category (Novel donor vs. Novel acceptor).
</details>

---

```{r echo = FALSE, fig.width=7.2, fig.height=3}
limit_bp = 30

modulo_group_sequence_df <- modulo_df %>%
  dplyr::mutate(sequence_type = paste(novel_type, sequence)) %>%
  dplyr::group_by(cluster, sequence_type, modulo) %>% 
  dplyr::summarise(n = n()) %>%
  dplyr::mutate(freq = (n / sum(n) * 100)) %>%
  dplyr::arrange(desc(sequence_type)) %>%
  dplyr::mutate(sequence_type = factor(sequence_type, levels = .$sequence_type %>% unique))

ggplot(modulo_group_sequence_df) +
  geom_bar(aes(x = modulo, y = freq, fill = cluster), stat = "identity", position = "dodge", color = "black", linewidth = 0.3) +
  scale_x_continuous(breaks = c(0, 1, 2), expand = c(0, 0.2)) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.05))) + 
  viridis::scale_fill_viridis(discrete = T, begin = 0.15, end = 0.75, name = "") +
  labs(x = "Modulo 3 of the distance", y = "Percentage of novel junctions") + 
  ggtitle(paste0("Mod.3 Percentages: Cluster, Splice site and Sequence"), subtitle = paste0("Tissue: ", tissue)) + 
  ggforce::facet_row(vars(sequence_type), labeller = labeller(sequence_type = sequence_labels)) +
  custom_gg_theme_subtitle + 
  theme(legend.key = element_rect(color = "black", linewidth = 0.10),
        axis.title.y = element_text(size = 9),
        plot.title = element_text(size = 13))
```

<details>
<summary>Expand details</summary>

* X-Axis: Distance in modulo 3 between the novel splice site and the reference splice site. Filtered by $distance\leq 100\;bp$.
* Y-Axis: Percentage of each modulo 3 value with respect to cluster (case and control), novel type (donor and acceptor) and sequence (intronic and exonic).
* Colour: Cluster type as case vs. control.
* Facets: Novel junction category (Novel donor vs. Novel acceptor) and sequence (intronic vs. exonic).
</details>

---

```{r results = "asis", echo = FALSE}
cat(addTitles(base_level, 1), " Delta MaxEntScan {.tabset -}\n", sep = "")
```

```{r echo=FALSE, message=F, warning = F, fig.height=4.5, fig.width=7.2}
delta_MES <- common_novel %>%
  tidyr::pivot_longer(c(delta_ss5score, delta_ss3score), names_to = "delta_type", values_to = "delta") %>%
  dplyr::filter(delta != 0) %>%
  dplyr::mutate(delta_type = ifelse(delta_type == "delta_ss5score", "delta MES 5'ss", "delta MES 3'ss")) %>%
  dplyr::mutate(delta_type = delta_type %>% as.factor()) %>%
  dplyr::mutate(delta_type = relevel(delta_type, ref = "delta MES 5'ss"))


ggplot(delta_MES) +
  geom_density(aes(x = delta, fill = cluster), alpha = 0.6, linewidth = 0.3, adjust = 1) +
  geom_vline(xintercept = 0) +
  scale_x_continuous(expand = c(0, 0)) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.05))) +
  viridis::scale_fill_viridis(discrete = T, begin = 0.15, end = 0.75, name = "") +
  labs(x = "Delta MaxEntScan score", y = "Density") +
  ggtitle(paste0("Delta MaxEntScan scores"), subtitle = paste0("Tissue: ", tissue)) +
  ggforce::facet_col(vars(delta_type), labeller = labeller(delta_type = delta_MES_labels)) +
  custom_gg_theme_subtitle + theme(legend.key = element_rect(color = "black", linewidth = .05))
```

<details>
<summary>Expand details</summary>

* X-Axis: Difference between the MaxEntScan score of the reference annotated intron and the novel junction.
* Y-Axis: Kernel density estimate (smoothed version of a histogram).
* Colour: Cluster type as case vs. control.
* Facets: Different splice sites.
</details>

---

```{r echo=FALSE, message=F, warning = F, fig.height=4.5, fig.width=7.2}
densities_5 <- delta_MES %>%
  dplyr::filter(delta_type == "delta MES 5'ss") %>%
  dplyr::group_by(cluster, delta_type) %>%
  dplyr::summarise(d = list(density(delta, from = min(.$delta), to = max(.$delta))))
densities_3 <- delta_MES %>%
  dplyr::filter(delta_type == "delta MES 3'ss") %>%
  dplyr::group_by(cluster, delta_type) %>%
  dplyr::summarise(d = list(density(delta, from = min(.$delta), to = max(.$delta))))

delta_5 <- tibble::tibble(x = densities_5$d[[1]]$x,
                          y = densities_5$d[[1]]$y - densities_5$d[[2]]$y)
delta_3 <- tibble::tibble(x = densities_3$d[[1]]$x,
                          y = densities_3$d[[1]]$y - densities_3$d[[2]]$y)
delta_df <- rbind(delta_5 %>% dplyr::mutate(delta_type = "delta MES 5'ss"),
                  delta_3 %>% dplyr::mutate(delta_type = "delta MES 3'ss")) %>%
  dplyr::arrange(desc(delta_type)) %>%
  dplyr::mutate(delta_type = factor(delta_type, levels = c("delta MES 5'ss", "delta MES 3'ss")))

ggplot(delta_df) +
  geom_area(aes(x = x, y = y, fill = "fill_color"), position = "identity") +
  scale_fill_manual(name = "", values = c("fill_color" = "#619cff"), labels = c("Case - Control"),
                    guide = guide_legend(override.aes = list(linewidth = .25))) +
  geom_line(aes(x = x, y = y)) +
  geom_vline(xintercept = 0) +
  geom_hline(yintercept = 0) +
  scale_x_continuous(expand = c(0, 0)) +
  scale_y_continuous(expand = expansion(mult = c(0.05, 0.05))) +
  labs(x = "Delta MaxEntScan score", y = "Density") +
  ggtitle(paste0("Delta MaxEntScan scores"), subtitle = paste0("Tissue: ", tissue)) +
  ggforce::facet_col(vars(delta_type), labeller = labeller(delta_type = delta_MES_labels)) +
  custom_gg_theme_subtitle + theme(legend.key = element_rect(color = "black", linewidth = 0.75))
```

<details>
<summary>Expand details</summary>

* X-Axis: Difference between the MaxEntScan score of the reference annotated intron and the novel junction.
* Y-Axis: Difference between the kernel density estimate for case samples and control samples.
* Colour: Cluster type as case vs. control.
* Facets: Different splice sites.
</details>

---

```{r results = "asis", echo = FALSE}
cat(addTitles(base_level, 1), " MSR {.tabset -}\n", sep = "")
```

```{r echo = F}
## Required library "scales"
suppressWarnings(suppressMessages(library(scales)))

plotMSR_2breaks <- function(MSR_df, MSR_site, MSR_density_max, title){
  ## Additional functions required for the plot
  round_any = function(x, accuracy, f = round){ f(x/accuracy) * accuracy }
  
  add_y_break <- function(plot, yval, n = 5) {
    p2 <- ggplot_build(plot)
    breaks <- p2$layout$panel_params[[1]]$y$breaks
    breaks <- breaks[!is.na(breaks)]
    
    plot +
      scale_y_continuous(expand = expansion(mult = c(0, 0.05)), 
                         breaks = round(sort(c(breaks, seq(0, yval, length.out = n)))))
  }
  
  ## Generate plot
  p <- ggplot(MSR_df) + 
    geom_density(aes(x = !!sym(MSR_site), fill = cluster), alpha = 0.5) +
    viridis::scale_fill_viridis(discrete = T, begin = 0.15, end = 0.75, name = "") +
    coord_trans(x = magnify_trans(intercept = 0.05, reducer = 50),
                y = magnify_trans(intercept = round_any(min(MSR_density_max), 100, f = ceiling), 
                                  reducer = 2*(max(MSR_density_max)-min(MSR_density_max))/min(MSR_density_max))) +
    geom_vline(xintercept = 0.05, linetype = 2, linewidth = 1) + 
    geom_hline(yintercept = round_any(min(MSR_density_max), 100, f = ceiling), linetype = 2, linewidth = 1) + 
    scale_x_continuous(breaks = c(0, 0.01, 0.02, 0.03, 0.04, 0.05, 0.2, 0.4, 0.6, 0.8, 1.0), 
                       expand = expansion(mult = c(0.0, 0.02)), limits = c(0, 1)) +
    labs(x = paste0("Mis-splicing ratio (", MSR_site, ")"), y = "Density") + 
    ggtitle(paste0("Mis-splicing ratio at ", title, " site"), subtitle = paste0("Tissue: ", tissue)) + 
    custom_gg_theme_subtitle + theme(legend.key = element_rect(color = "black", linewidth = .05))
  
  ## Add Y-Axis discontinuity
  add_y_break(p, round_any(min(MSR_density_max), 100, f = ceiling))
}

plotMSR_1break <- function(MSR_df, MSR_site, title){
  ggplot(MSR_df) + 
    geom_density(aes(x = !!sym(MSR_site), fill = cluster), alpha = 0.5) +
    viridis::scale_fill_viridis(discrete = T, begin = 0.15, end = 0.75, name = "") +
    scale_y_continuous(expand = expansion(mult = c(0, 0.05))) + 
    coord_trans(x = magnify_trans(intercept = 0.05, reducer = 50)) +
    geom_vline(aes(xintercept = 0.05), linetype = 2, linewidth = 1) + 
    scale_x_continuous(breaks = c(0, 0.01, 0.02, 0.03, 0.04, 0.05, 0.2, 0.4, 0.6, 0.8, 1.0), 
                       expand = expansion(mult = c(0.0, 0.02)), limits = c(0, 1)) +
    labs(x = paste0("Mis-splicing ratio (", MSR_site, ")"), y = "Density") + 
    ggtitle(paste0("Mis-splicing ratio at ", title, " site"), subtitle = paste0("Tissue: ", tissue)) + 
    custom_gg_theme_subtitle + theme(legend.key = element_rect(color = "black", linewidth = .05))
}

magnify_trans <- function(intercept, reducer) {
  trans <- function(x, i = intercept, r = reducer) {
    sapply(x, function(x) {
      if (x < i) x
      else x / r + i
    }) %>% as.numeric()
  }
  
  inv <- function(x, i = intercept, r = reducer) {
    sapply(x, function(x) {
      if(!is.na(x)) {
        if (x < i) x
        else (x - i) * r
      }
    }) %>% as.numeric()
  }
  trans_new(name = 'custom',
            transform = trans,
            inverse = inv
  )
}
```

```{r echo = F, message=F, warning = F, fig.height=4.5, fig.width=7.2}
MSR_df <- common_introns
MSR_densities <- MSR_df %>%
  tidyr::pivot_longer(cols = c("MSR_Donor", "MSR_Acceptor"), names_to = "MSR_type", values_to = "MSR") %>%
  dplyr::group_by(MSR_type, cluster) %>% 
  dplyr::summarise(d = list(density(MSR, from = min(.$MSR), to = max(.$MSR)))) %>%
  dplyr::rowwise() %>%
  dplyr::mutate(max = d$y %>% max) %>%
  dplyr::ungroup()

for(MSR_site in c("MSR_Acceptor", "MSR_Donor")){
  MSR_density_max <- MSR_densities %>% dplyr::filter(MSR_type == MSR_site) %>% dplyr::pull(max)
  title = ifelse(MSR_site == "MSR_Donor", "donor", "acceptor")
  
  ## If the difference between the two maximums is too high, we plot with an axis break
  if(max(MSR_density_max)/min(MSR_density_max) > 20){
    plotMSR_2breaks(MSR_df, MSR_site, MSR_density_max, title) %>% print()
  }else{
    plotMSR_1break(MSR_df, MSR_site, title) %>% print()
  }
  
  ## Add white spaces between the plots
  if(MSR_site == "MSR_Acceptor") cat("<br><br><br>")
}
```

<details>
<summary>Expand details</summary>

* X-Axis: The mis-splicing ratio.
* Y-Axis: Kernel density estimate.
* Colour: Cluster type as case vs. control.
* Facets: Different splice sites.
* Notes: Distributions have a discontinuity at X = 0.05 to better represent all possible MSR values. In some cases, a discontinuity in the Y-axis is also present if the difference between the two sample types exceeds a predefined threshold.
</details>

---

```{r results = "asis", echo = FALSE}
cat(addTitles(base_level, 1), " Stats {-}\n", sep = "")
```

```{r results = "asis", echo = FALSE}
cat(addTitles(base_level, 2), " <ins>Unique annotated introns</ins> {-}\n", sep = "")
```

```{r echo = F}
common_introns %>%
  dplyr::group_by(ref_type, cluster) %>%
  dplyr::count(N = n()/2) %>%
  dplyr::mutate("percent" = round((n/N)*100, digits = 2)) %>%
  `colnames<-`(c("Mis-spliced site", "Cluster", "# Annotated introns", "# Ann. introns by splice site", "Percent [%]")) %>%
  kableExtra::kbl(booktabs = T, linesep = "") %>%
  kableExtra::kable_classic(full_width = T, "hover", "striped", html_font = "Cambria", font_size = 14) %>%
  kableExtra::row_spec(0, bold = T, font_size = 16)
```

<details>
<summary>Expand details</summary>

* \# Annotated introns: number of total annotated introns in the cluster. Since we limit the annotated introns to those common across all samples and clusters, the number is the same for both clusters.
* \# Ann. introns by splice site: the number of annotated introns for the cluster and splice site.
* Percent [%]: the percentage that the annotated introns in the splice site represent to the total number of introns in the cluster:

$$
\text{Percent [%]}=\frac{\text{# Ann. introns by splice site}}{\text{# Annotated introns}}*100\%
$$
</details>

---

```{r results = "asis", echo = FALSE}
cat(addTitles(base_level, 2), " <ins>Reads - annotated introns</ins> {-}\n", sep = "")
```

```{r echo = F}
common_introns %>%
  dplyr::group_by(cluster) %>%
  dplyr::mutate(total_sum_counts = sum(ref_reads)) %>%
  dplyr::select(cluster, total_sum_counts) %>%
  dplyr::group_by(cluster, total_sum_counts) %>%
  dplyr::count() %>%
  dplyr::left_join(metadata %>% select(Type, total_reads),
                   by = c("cluster" = "Type"), multiple = "all") %>%
  dplyr::group_by(cluster, total_sum_counts, n) %>%
  dplyr::summarise(total_read_depth = sum(total_reads)) %>%
  dplyr::relocate(n, total_read_depth, .after = cluster) %>%
  dplyr::mutate(percent = round(total_sum_counts/total_read_depth*100, digits = 2)) %>%
  `colnames<-`(c("Cluster", "# Annotated introns", "# Reads in cluster", "# Ann. intron reads", "Percentage [%]")) %>%
  kableExtra::kbl(booktabs = T, linesep = "") %>%
  kableExtra::kable_classic(full_width = T, "hover", "striped", html_font = "Cambria", font_size = 14) %>%
  kableExtra::row_spec(0, bold = T, font_size = 16)
```

<details>
<summary>Expand details</summary>

* \# Annotated introns: number of common annotated introns in the cluster. Since we limit the annotated introns to those common across all samples and clusters, the number is the same for both clusters.
* \# Reads in cluster: total number of reads from the BAM files that constitutes the cluster.
* \# Ann. intron reads: reads associated to the annotated introns in the cluster. Only common introns are being counted.
* Percent [%]: the percentage that the annotated intron reads represent to the total number of reads in the cluster.

$$
\text{Percent [%]}=\frac{\text{# Ann. intron reads}}{\text{# Reads in cluster}}*100\%
$$
</details>

---

```{r results = "asis", echo = FALSE}
cat(addTitles(base_level, 2), " <ins>Reads - novel junctions</ins> {-}\n", sep = "")
```

```{r echo = F}
common_novel %>%
  dplyr::group_by(novel_type, cluster) %>%
  dplyr::mutate(total_sum_counts = sum(novel_reads)) %>%
  dplyr::select(novel_type, cluster, total_sum_counts) %>%
  dplyr::left_join(metadata %>% 
                     dplyr::select(Type, total_reads) %>%
                     dplyr::group_by(Type) %>%
                     dplyr::summarise(total_reads = sum(total_reads)),
                   by = c("cluster" = "Type"), multiple = "all") %>%
  dplyr::mutate(percent = round(total_sum_counts/total_reads*100, digits = 2)) %>%
  dplyr::group_by(cluster, novel_type, total_sum_counts, total_reads, percent) %>%
  dplyr::summarize(n = n()) %>%
  dplyr::group_by(cluster) %>%
  dplyr::mutate(N = sum(n)) %>%
  dplyr::relocate(N, n, total_reads, .after = novel_type) %>%
  `colnames<-`(c("Cluster", "Novel type", "# Novel junctions", "# Novel junc. in splice site", 
                 "# Reads in cluster", "# Novel junc. reads", "Percentage [%]")) %>%
  kableExtra::kbl(booktabs = T, linesep = "") %>%
  kableExtra::kable_classic(full_width = T, "hover", "striped", html_font = "Cambria", font_size = 14) %>%
  kableExtra::row_spec(0, bold = T, font_size = 16)
```

<details>
<summary>Expand details</summary>

* \# Novel junctions: total number of novel junctions for that particular cluster.
* \# Novel junc. in splice site: total number of novel junctions for that particular splice site in the cluster.
* \# Reads in cluster: total number of reads from the BAM files that constitutes the cluster.
* \# Novel junc. reads: reads associated to the novel junctions in the cluster and splice site.
* Percent [%]: the percentage that the novel junction reads represent to the total number of reads in the cluster.

$$
\text{Percent [%]}=\frac{\text{# Novel junc. reads}}{\text{# Reads in cluster}}*100\%
$$
</details>